\documentclass[16pt]{article}
\usepackage[affil-it]{authblk}
\title{Integrating Static Analysis and Large Language Models for Effective Code Comprehension}
\author{George Fakidis}
\affil{Athens University of Economics and business \\ Department of Computer Science}
\date{}

\begin{document}

    \maketitle

    \large
    In contemporary software engineering, most time is spent reading code instead of writing code
    due to maintenance and enhancement efforts in existing codebases and more generally the iterative nature of software development.
    Therefore, it is economically imperative that those professionals are given the tools to enhance their effectiveness in reading and understanding code.

    Code is structured text and AI and LLMs have succeeded massively in helping users
    understand pieces of unstructured text by providing summaries, explanations without assuming
    previous knowledge and asking questions based on the provided documents.
    This indicates that AI and LLMs could be a good fit for streamlining the process of reading and understanding code.

    Static Program Analysis is the machine-automated process of reading and understanding
    code using sophisticated rules and methods that allow the extraction of insights not immediately
    evident to the software professional and the presentation of different perspectives in order to
    more deeply understand the code at hand.

    By combining the two complementary techniques into one coherent agent,
    explaining code through LLM techniques utilizing Static Analysis tools,
    we can tailor the explanations to the specific problem the developer is currently facing.
    It is important to note that existing tooling and approaches overly focus on writing new code fast
    instead of understanding existing code efficiently.

    More specifically, we intend to develop a methodological framework for integrating
    Static Analysis and LLMs taking into consideration their respective strengths and
    weaknesses in order to achieve significant increases in developer effectiveness and productivity.

\end{document}